Learning and Evaluation of \strips\ Action Models with Classical Planning

### 2018-05-02 doc/aij18/aij18.tex
appendix added to aij18 paper
updates added to sections 5 and 6 AIJ18 paper
updates to the aij18 paper

\begin{abstract}
  This paper presents a novel approach for learning \strips\ action models from observations of plan executions that compiles this learning task into classical planning. The compilation approach is flexible to various amount and kind of available input knowledge; learning examples can range from plans (with their corresponding initial state) to sequences of state observations or even just a set of initial and final states (where no intermediate action/state is known). The compilation accepts also partially specified action models and can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified. What is more, the compilation is extensible to assess how well a given \strips\ action model matches an observation of a plan execution. This extension is relevant since it allows us to evaluate the quality of the learned models with respect to the true models but also with respect to test sets of observations of plan executions, which is useful when the actual models are not available. The empirical performance of our compilation approach is evaluated learning action models for a wide range of classical planning domains from the International Planning Competition (IPC) and using these two evaluation approaches.
\end{abstract}

@@

>>>>

\begin{abstract}
  This paper presents a novel approach for learning \strips\ action models from observations of plan executions that compiles this learning task into classical planning. The compilation approach is flexible to various amount and kind of available input knowledge; learning examples can range from plans (with their corresponding initial state) to sequences of state observations or even just a set of initial and final states (where no intermediate action or state is known). The compilation accepts also partially specified action models and can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified. What is more, the compilation is extensible to assess how well a given \strips\ action model matches the observation of a plan execution. This extension allows us to evaluate the quality of the learned models with respect to the actual models but also, with respect to observations of plan executions that act as a {\em test set}, which is useful when the actual models are not available. The performance of our compilation approach is evaluated learning action models for a wide range of classical planning domains from the International Planning Competition (IPC) and using these two evaluation approaches.
\end{abstract}

@@

\input{appendix}

### 2018-05-02 doc/aij18/introduction.tex
updates to the aij18 paper

Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that can be defined as a classical planning compilation. A solution to the classical planning task that results from our compilation is a sequence of actions that determines the preconditions and effects of a \strips\ model such that this \strips\ model satisfies the observed plan executions that are given as input.

The compilation approach is appealing by itself because it leverages off-the-shelf planners for learning and because its practicality allow us to report model learning results over a wide range of IPC planning domains. Moreover, it opens up a way towards \emph{bootstrapping} planning action models, enabling a planner to gradually learn/update its action model. Apart from these, our classical planning compilation presents the following contributions:
\begin{enumerate}
\item {\em Input Flexibility}. The compilation is flexible to various amount and kind of available input knowledge. Learning examples can range from a set of plans (with their corresponding initial state) or state observations, to just a set of initial and final states where no intermediate action or state is observed. This fact makes the compilation robust to partially obervability of the intermediate input data. Last but not least, the compilation also accepts input knowledge about the structure of the actions in the form of partially specified action models, where some preconditions and effects are a priori known. % Also accepts background knowledge

\item {\em Model Validation}. The compilation poses a general framework to assess the validation of a generative model (provided that it is STRIPS compilable) with a given set of observations. This validation capacity goes beyond the functionality of VAL~\cite{howey2004val} since we do not require a full plan or a full action model.

\item {\em Model Evaluation}. The compilation can assess how well a given \strips\ action model matches a {\em test set} with observations of plan executions. This allows us to assess learning performance without having the actual model and again, evaluation with our compilation is flexible to various amount and kind of available data in the given {\em test set}. 
\end{enumerate}

A first description of the compilation previously appeared in the conference paper~\cite{aineto2018learning}. Compared to the conference paper, this work includes the following novel material:
\begin{itemize}
\item The formulation of {\em model evaluation} as the task of assessing the amount of edition required by the input action model to induce given observations of plan execution that serve as a test set.
\item A unified formulation for learning and evaluating \strips\ action models from observed executions of plans.
\item Leveraging {\em background knowledge} (given as planning constraints either in the form of {\em state constraints} or {\em trajectory constraints}) for learning/evaluating/recognizing \strips\ action models.
\item A more complete empirical evaluation of the compilation approach. Our evaluation analyses how the amount of input knowledge affects to the performance of the compilation approach.
\end{itemize}

Section~\ref{sec:Section2} reviews related work on learning planning action models. Section~\ref{sec:Section3} introduces the classical planning model with {\em conditional effects} (a requirement of the proposed compilation) and the \strips\ action model (the target of our learning and evaluation tasks). Section~\ref{sec:Section4} formalizes the learning of \strips\ action models with regard to different amount and kind of available input knowledge. Sections~\ref{sec:Section5} and ~\ref{sec:Section6} describe our compilation approach for addressing the formalized learning tasks its extension to evaluate and recognize \strips\ action models. Section~\ref{sec:Section7} reports the data collected in a two-fold empirical evaluation of our learning approach: First, the learned \strips\ action models are tested with observations of plan executions and second, the learned models are compared to the actual models. Finally, Section ~\ref{sec:Section8} discusses the strengths and weaknesses of the compilation approach and proposes several opportunities for future research.

>>>>

Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that can be defined as a classical planning compilation. A solution to the classical planning task that results from our compilation is a sequence of actions that determines the preconditions and effects of a \strips\ model such that this \strips\ model satisfies a set of plan executions that is given as input.

The compilation approach is appealing by itself because it leverages off-the-shelf planners for learning and because its practicality allow us to report model learning results over a wide range of IPC planning domains. Moreover, it opens up a way towards \emph{bootstrapping} planning action models, enabling a planner to gradually learn/update its action model. Apart from these, our classical planning compilation presents the following contributions:
\begin{enumerate}
\item {\em Input flexibility}. The compilation is flexible to various amount and kind of available input knowledge. Learning examples can range from a set of plans (with their corresponding initial state) or state observations, to just a set of initial and final states where no intermediate action or state is observed. This fact makes the compilation robust to the partial obervability of the intermediate input data. Last but not least, the compilation also accepts input knowledge about the structure of the actions in the form of partially specified action models, that is learning action models when some preconditions and effects are a priori known. % Also accepts background knowledge

\item {\em Model validation}. The compilation poses a novel framework to assess the validation of a STRIPS model with respect to a given set of observations. This validation capacity goes beyond the functionality of VAL~\cite{howey2004val} since it does not require a full plan or a full action model to determine validation. The relevance of this validation framework overcomes classical planning because of the wide range of different generative models that are STRIPS compilable (STRIPS is {\em Turing complete}).

\item {\em Model evaluation}. The compilation can assess how well a given \strips\ action model matches a {\em test set} with observations of plan executions. The idea is to assess the amount of edition required by the input action model to induce given observations of plan execution that serve as a test set. This allows us to assess learning performance without having the actual model and again, evaluation with our compilation is flexible to various amount and kind of available data in the given {\em test set}. 
\end{enumerate}

A first description of the compilation previously appeared in the conference paper~\cite{aineto2018learning}. Compared to that conference paper, this work includes the following novel material:
\begin{itemize}
\item A unified formulation for learning and evaluating \strips\ action models from observed executions of plans. 
\item Leveraging diverse forms of {\em background knowledge} for learning and evaluating \strips\ action models.
\item A complete empirical evaluation of the compilation approach. Our evaluation analyses how the amount of input knowledge affects to the performance of the compilation approach when learning and evaluating \strips\ action models.
\end{itemize}

Section~\ref{sec:Section2} reviews related work on learning planning action models. Section~\ref{sec:Section3} introduces the classical planning model and the \strips\ action model (the target of our learning and evaluation tasks). Section~\ref{sec:Section4} formalizes the learning of \strips\ action models with regard to different amount and kind of available input knowledge. Sections~\ref{sec:Section5} and ~\ref{sec:Section6} describe our compilation approach for addressing the formalized learning tasks its extension to evaluate and recognize \strips\ action models. Section~\ref{sec:Section7} reports the data collected in a two-fold empirical evaluation of our learning approach: First, the learned \strips\ action models are tested with observations of plan executions and second, the learned models are compared to the actual models. Finally, Section ~\ref{sec:Section8} discusses the strengths and weaknesses of the compilation approach and proposes several opportunities for future research.
